### 2015-05-08 whoop.tex
A bit of progress

\ADComment{Please ignore my comments in this section for now.}

We now present \emph{static pairwise lockset analysis}, a novel method for data race analysis in concurrent device drivers.
In Section~\ref{ } we explain conceptually how the approach works.
In Section~\ref{ } we describe our implementation of static pairwise lockset analysis in a practical tool,
Whoop, that can be applied directly to driver source code, to check whether pairs of driver entry points are 
free from data races.

We demonstrate in Section~\ref{ } both that the Whoop technique and tool has value in its own right as a stand-alone static analyser, and that results from Whoop can be exploited to significantly boost performance of a more precise static analysis for concurrency offered by the CORRAL tool, based on context bounding and stratified inlining.

\subsection{Static pairwise lockset analysis}

\ADComment{Ally: first, let us try to explain really clearly in words how the technique works.}

\ADComment{Second, let's introduce some notation to describe the thing more formally.}

We present our approach in a semi-formal manner, in the context of the following simple concurrent programming model.  For ease of presentation only, the model does not features aggregate data types, pointers or dynamic memory allocation.  These \emph{are} handled by our implementation, and in \ADComment{TODO} we discuss interesting practical issues arising from the handling of a full-blown language.  \ADComment{I want to say that we do not treat dynamic thread creation because driver threads do not create further driver threads.  However, I am not sure (a) whether this is true, nor (b) whether even if it is true it would be a problem for us.}

@@

We consider a concurrent programming model in which an unbounded number of threads execute a set of pre-defined \emph{thread templates}.  In the context of device drivers, a thread template corresponds to a driver entry point.  At any given point of execution a certain number of threads are active, each thread executing a particular template.  As in the case of drivers, where multiple distinct entry points as well as multiple instances of the same entry point may run concurrently, two distinct threads in our model may execute according to the same template or according to different templates.  Further threads may be launched at any point during execution, just as in the context of device drivers where the OS may invoke additional driver entry points.

@@

\subsection{Implementation in Whoop}

>>>>

We now present \emph{symbolic pairwise lockset analysis}, a novel method
for data race analysis in concurrent device drivers.  In
Section~\ref{sec:symbolicpairwise} we describe how the approach works
in a semi-formal manner, with respect to a simple concurrent
programming model.  In Section~\ref{sec:whoopimplementation} we
explain how we have implemented symbolic pairwise lockset analysis in
a practical tool, Whoop, that can be applied directly to driver source
code, to check whether pairs of driver entry points can race with one
another.

Our experimental evaluation (Section~\ref{evaluation}) demonstrates that the Whoop technique has value as a stand-alone analyser, and that results from Whoop can be exploited to significantly boost performance of a more precise symbolic analysis for concurrency offered by the CORRAL tool, based on context bounding and stratified inlining; we discuss the latter approach in Section~\ref{whoop:bugfinding}.

\subsection{Symbolic pairwise lockset analysis}\label{sec:symbolicpairwise}

In short, symbolic pairwise lockset analysis works by considering, for a given device driver, every pair of entry points that can potentially execute concurrently.  For each such pair, we use symbolic verification to check whether it is possible for the pair to race, accounting for the effects of additional entry points by treating the driver shared state abstractly so that when an entry point reads from the shared state, a nondeterministic value is returned.  Symbolic verification of an entry point pair works by instrumenting each entry point with additional state to record locksets (see Section~\ref{bg:lockset}), and attempting to verify a sequential program that executes the instrumented entry points in sequence, and subsequently asserts that the lockset for each shared variable is non-empty.  Verification of the resulting sequential program can be undertaken using any sound method; in practice we employ the Boogie verification engine, which requires procedure specifications and loop invariants to be generated, after which verification conditions are generated and discharged to an automated theorem prover.

We now detail the approach in a semi-formal manner, in the context of a simple concurrent programming model.

@@

We consider a concurrent programming model in which an unbounded number of threads execute a set of pre-defined \emph{thread templates}.  In the context of device drivers, a thread template corresponds to a driver entry point.  At any given point of execution a certain number of threads are active, each thread executing a particular template.  As in the case of drivers, where multiple distinct entry points as well as multiple instances of the same entry point may run concurrently, two distinct threads in our model may execute according to the same template or according to different templates.  Further threads may be launched at any point during execution, just as in the context of device drivers where the OS may invoke additional driver entry points.  For ease of presentation only, our model does not feature aggregate data types, pointers or dynamic memory allocation.  These \emph{are} handled by our implementation, and in Section~\ref{sec:whoopimplementation} we discuss interesting practical issues arising from the handling of a full-blown language.

@@

\subsection{Implementation in Whoop}\label{sec:whoopimplementation}
